Numerical simulation of the Cordilleran ice sheet \\ through the last glacial cycle

### 2014-06-24 kappa.tex
Use cleveref package for homogeneized int. refs.

% Base class and packages
\documentclass{article}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{lineno}
\usepackage{natbib}
\usepackage[onehalfspacing]{setspace}
\usepackage{physics}
\usepackage[colorlinks, citecolor=blue]{hyperref}
%\usepackage{graphicx}

@@

% Internal references
\newcommand{\eqn}[1]{Eqn.~\ref{eqn:#1}}
\newcommand{\fig}[1]{Fig.~\ref{fig:#1}}
\newcommand{\tab}[1]{Table~\ref{tab:#1}}

@@

Letting ${\doteps_e}^2 = \frac{1}{2} \tr(\SRT^2)$ and
$B=A^{-1/n}$, the constitutive law (\eqn{glenslaw}) can be rewritten using a
sometimes more convenient, inverted formulation,

@@

where $\vect{q}$ represents the heat flow and
${\frac{\nu\dot{\epsilon_e}^2}{4} = \tr(\DST\SRT)}$ is a
source term corresponding to strain heat release. In the case where ice
temperature, $T$, is below the pressure-melting point, the enthalpy can be
expressed as $H=c\,T$, while the heat flow is given by Fourier's law,
$\vect{q} = k\,\grad{T}$. Consequently, the enthalpy equation (\eqn{enthalpy})
can be rewritten using a more familiar, cold-ice, temperature formulation,

@@

where $c$ denotes the specific heat capacity of the ice and $k$ is its thermal
conductivity. Equations~\ref{eqn:incompressibility}, \ref{eqn:stressbalance},
\ref{eqn:glenslaw}, and \ref{eqn:enthalpy} constitute the thermodynamic basis
of our model. However, their resolution in full form implies a computational
demand too high for multi-millennial, continental-scale applications, such as
the numerical modelling of the Cordilleran ice sheet through the last glacial
cycle, thus requiring shallow approximations.

@@

Distinguishing the hydrostatic and deviatoric components of the stress tensor,
the balance of stresses (\eqn{stressbalance}) can be expanded into

\begin{equation}
    \label{eqn:stressbalance}
    \tdiv{\DST} - \grad{p} + \rho\,\vect{g} = \vect{0}.
\end{equation}

@@

Because $\vsia$ is small where $\vssa$ is large and
reciprocally, we can expect the model to perform well in regions where
conditions for applicability of either the SIA or the SSA are met. However,
equation~\ref{eqn:siassa} is a heuristic, and its validity in the zone of
transition between shear flow and longitudinal flow has not yet been
mathematically proven.

@@

where $\vect{\tau}_b$ is the basal traction force, $\vv_b$ is the basal
velocity, and $v_{th}$ is an arbitrary velocity threshold. In the case of
$q=0$, equation~\ref{eqn:pseudoplastic} corresponds to a purely plastic law,

@@

The modelled ice sheet interacts with the underlying bedrock in two ways.
Firstly, bedrock temperatures are computed to a given depth by appending the
ice enthalpy model \eqn{enthalpy} with an underlying bedrock thermal model.
The only process accounted for is heat conduction, hence this model consists of
a simple diffusion equation.

@@

where $\nu_m$ is mantle viscosity, $\rho_l$ is lithosphere density, $D$ is the
lithosphere's flexural rigidity, and $\sigma_{zzb}$ corresponds to the ice load
\citep{Bueler.etal.2007}. Here, $\Delta^2$ designs the biharmonic (Laplacian
square) operator of
linear elastic theory, while $|\grad|$ is a pseudo-differential operator
defined through the Fourier transform by \citet[Eqn.~6]{Bueler.etal.2007}. In
the left-hand part of this equation, the first term accounts for mantle
relaxation, the second for point-wise isostasy, and the third for elastic
flexure. Due to high mantle viscosity $\nu_m$, there exists a time lag between
the ice sheet growth and the isostatic bedrock response.
%$\nu=1\times10^{21}$\,Pa\,s
%$D=5\times10^{24}$\,N\,m
%$\rho_r = 3300$\,kg\,m$^{-3}$

>>>>

% Base class and packages
\documentclass{article}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{lineno}
\usepackage{natbib}
\usepackage[onehalfspacing]{setspace}
\usepackage{physics}
\usepackage[colorlinks, citecolor=blue]{hyperref}
\usepackage[sort]{cleveref}  % needs to be loaded after hyperref
%\usepackage{graphicx}

% Copernicus-style internal references
\crefname{chapter}{Chap.}{Chaps.}
\crefname{equation}{Eq.}{Eqs.}
\crefname{figure}{Fig.}{Figs.}
\crefname{section}{Sect.}{Sects.}
\crefname{table}{Table}{Tables}

@@

@@

Letting ${\doteps_e}^2 = \frac{1}{2} \tr(\SRT^2)$ and $B=A^{-1/n}$, the
constitutive law (\ref{eqn:glenslaw}) can be rewritten using a sometimes more
convenient, inverted formulation,

@@

where $\vect{q}$ represents the heat flow and
${\frac{\nu\dot{\epsilon_e}^2}{4} = \tr(\DST\SRT)}$ is a
source term corresponding to strain heat release. In the case where ice
temperature, $T$, is below the pressure-melting point, the enthalpy can be
expressed as $H=c\,T$, while the heat flow is given by Fourier's law,
$\vect{q} = k\,\grad{T}$. Consequently, the enthalpy equation
(\ref{eqn:enthalpy}) can be rewritten using a more familiar, cold-ice,
temperature formulation,

@@

where $c$ denotes the specific heat capacity of the ice and $k$ is its thermal
conductivity. \Cref{eqn:incompressibility,eqn:stressbalance,eqn:glenslaw,%
eqn:enthalpy} constitute the thermodynamic basis
of our model. However, their resolution in full form implies a computational
demand too high for multi-millennial, continental-scale applications, such as
the numerical modelling of the Cordilleran ice sheet through the last glacial
cycle, thus requiring shallow approximations.

@@

Distinguishing the hydrostatic and deviatoric components of the stress tensor,
the balance of stresses (\ref{eqn:stressbalance}) can be expanded into

\begin{equation}
    \tdiv{\DST} - \grad{p} + \rho\,\vect{g} = \vect{0}.
\end{equation}

@@

Because $\vsia$ is small where $\vssa$ is large and
reciprocally, we can expect the model to perform well in regions where
conditions for applicability of either the SIA or the SSA are met. However,
\cref{eqn:siassa} is a heuristic, and its validity in the zone of
transition between shear flow and longitudinal flow has not yet been
mathematically proven.

@@

where $\vect{\tau}_b$ is the basal traction force, $\vv_b$ is the basal
velocity, and $v_{th}$ is an arbitrary velocity threshold. In the case of
$q=0$, \cref{eqn:pseudoplastic} corresponds to a purely plastic law,

@@

The modelled ice sheet interacts with the underlying bedrock in two ways.
Firstly, bedrock temperatures are computed to a given depth by appending the
ice enthalpy model (\ref{eqn:enthalpy}) with an underlying bedrock thermal
model. The only process accounted for is heat conduction, hence this model
consists of a simple diffusion equation.

@@

where $\nu_m$ is mantle viscosity, $\rho_l$ is lithosphere density, $D$ is the
lithosphere's flexural rigidity, and $\sigma_{zzb}$ corresponds to the ice load
\citep{Bueler.etal.2007}. Here, $\Delta^2$ designs the biharmonic (Laplacian
square) operator of
linear elastic theory, while $|\grad|$ is a pseudo-differential operator
defined through the Fourier transform by \citet[Eq.~6]{Bueler.etal.2007}. In
the left-hand part of this equation, the first term accounts for mantle
relaxation, the second for point-wise isostasy, and the third for elastic
flexure. Due to high mantle viscosity $\nu_m$, there exists a time lag between
the ice sheet growth and the isostatic bedrock response.
%$\nu=1\times10^{21}$\,Pa\,s
%$D=5\times10^{24}$\,N\,m
%$\rho_r = 3300$\,kg\,m$^{-3}$
